---
title: "Cross-validation"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Simulate data

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )
```

Look at data

```{r}
nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

## Cross-validation by hand

Get training and testing datasets

```{r}
train_df = sample_n(nonlin_df, size = 80)
test_df = anti_join(nonlin_df, train_df, by = "id")
```

Fit three models.

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = gam(y ~ s(x), data = train_df)
wiggle_mod = gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

Can I see what I just did ... ?

```{r}
train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df %>% 
  gather_predictions(linear_mod, smooth_mod, wiggle_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() +
  geom_line(aes(y = pred), color = "red") +
  facet_grid(. ~ model)
```

Look at prediction accuracy.

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggle_mod, test_df)
```

## Cross validation using 'modelr'

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100)
```

What is happening here ...

```{r}
cv_df %>% 
  pull(train) %>% 
  .[[1]] %>% 
  as_tibble()

cv_df %>% 
  pull(test) %>% 
  .[[1]] %>% 
  as_tibble()
```

```{r}
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Let's try to fit models and get RMSEs for them.

```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_mod = map(.x = train, ~lm(y ~ x, data = .x)),
    smooth_mod = map(.x = train, ~gam(y ~ s(x), data = .x)),
    wiggle_mod = map(.x = train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(.x = smooth_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_wiggle = map2_dbl(.x = wiggle_mod, .y = test, ~rmse(model = .x, data = .y))
  )
```

What do these results say about model choices?

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Compute averages ... 

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  group_by(model) %>% 
  summarize(avg_rmse = mean(rmse))
```

## Try on a real dataset ... 

Import data.

```{r}
child_growth = read_csv("./data/nepalese_children.csv") %>% 
  mutate(
    weight_cp = (weight > 7) * (weight - 7)
  )
```

Weight vs arm circumference.

```{r}
child_growth %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.3)
```

Fit the models I care about.

```{r}
linear_mod = lm(armc ~ weight, data = child_growth)
pwlin_mod = lm(armc ~ weight + weight_cp, data = child_growth)
smooth_mod = gam(armc ~ s(weight), data = child_growth)
```

```{r}
child_growth %>% 
  gather_predictions(linear_mod, pwlin_mod, smooth_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.3) +
  geom_line(aes(y = pred), color = "red") +
  facet_grid(~model)
```

Try to understand model fit using cross validation.

```{r}
cv_df = 
  crossv_mc(child_growth, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

See if I can fit the models to the splits ...

```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_mod  = map(train, ~lm(armc ~ weight, data = .x)),
    pwl_mod     = map(train, ~lm(armc ~ weight + weight_cp, data = .x)),
    smooth_mod  = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_pwl    = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)))
```

Violin plot of RMSEs

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
